### 2020-09-16 README.Rmd
deal with resprout issue
Updated upto the blockCV stuff.
readme through shuffle RMSE comparison

@@

We use the function `add_buffer_variable` to identifiy which trees are inside of a buffer region of size `max_dist`.

@@

@@

### Model specification

Next we specify the growth model we want. We will model growth as a function of tree identity, size (in dbh), and the identity and size (in dbh) of its neighbors. The model can be run on different grouping of individuals (e.g., based on species or trait groupings). It can also be run with different notions of competition:

* `model_number = 1`: No competition growth only depends on focal tree `dbh` and grouping.
* `model_number = 2`: Competition but identity of competitor does not matter, just sum of competitior basal area.
* `model_number = 3`: Full model with competition including competitor identity.

```{r, eval = run_bw}
# Big Woods
bw_specs <- bw_growth_df_orig %>% 
  get_model_specs(model_number = 3, species_notion = 'trait_group')
bw_specs
```

@@

This requires the `growth_df` data frame, `max_dist` scalar defining competitive range, `cv_fold_size` defining the size of the spatial cross-validation blocks, and `model_specs` as inputs.

```{r, eval = run_bw}
# Big Woods
if (!file.exists("focal_vs_comp_bw.Rdata")) {
  tic()
  focal_vs_comp_bw <- bw_growth_df_orig %>% 
    create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
  toc()
  save(focal_vs_comp_bw, file = "focal_vs_comp_bw.Rdata")
} else {
  load("focal_vs_comp_bw.Rdata")
}
```

@@

Now we are ready to fit the competition model with `fit_bayesian_model` this takes two inputs: `focal_vs_comp` the data frame with focal trees connected to their competitors and `model_specs` which specifies the notion of competition.

```{r, eval = run_bw}
# Big Woods
tic()
bw_fit_model <- focal_vs_comp_bw %>% 
  fit_bayesian_model(model_specs = bw_specs)
toc()
```

This output has the posterior parameters for the specified competition model. If `model_number = 1` this includes just the beta values for how focal tree DBH and identity affect growth. If `model_number = 2` it also includes a vector of lambdas for how total competitor basal area affect individuals of each group. If `model_number = 3` then lambda is a matrix which gives how individuals of each focal group respond to the competition of each competitor group. This `posterior_param` output can be used to get predicted growths for each individual (with `predict_bayesain_model`) to test how well the model performs. Or this `posterior_param` output can be plots (either the betas or lambdas) to understand what controls individual growth.

```{r, eval = run_bw}
# Big Woods
bw_growth_df <- focal_vs_comp_bw %>% 
  predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model) %>% 
  right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID"))

bw_growth_df <- bw_growth_df %>% 
  st_as_sf()

# Observed vs predicted growth  
ggplot(bw_growth_df, aes(x = growth, y = growth_hat)) +
  geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
  stat_smooth(method = 'lm') +
  geom_abline(slope = 1, intercept = 0) +
  coord_fixed() + 
  labs(
    x = "Observed growth in dbh", y = "Predicted growth in dbh", 
    title = "Predicted vs Observed Growth"
  )

reslab <- expression(paste('Residual (cm ',y^{-1},')'))
bw_growth_df %>% 
  # Need to investigate missingness
  filter(!is.na(growth_hat)) %>% 
  mutate(
    error = growth - growth_hat,
    error_bin = cut_number(error, n = 5), 
    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
  ) %>% 
  ggplot() + 
  geom_sf(aes(col = error_compress), size = 0.4) + 
  theme_bw() + 
  scale_color_gradient2(
    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf", 
    name = reslab,
    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
    labels = c('< -0.75', '-0.5', '0.25', '0', '0.25', '0.5', '> 0.75')) +
  labs(x = "Meter", y = "Meter")
```

>>>>

# we need to filter out the resprouts
bw_census_2014 <- bw_census_2014 %>% 
  filter(!str_detect(codes, 'R'))

@@

We use the function `add_buffer_variable` to identifiy which trees are inside of a buffer region of size `max_dist`. The user will need to specify a study region and convert to a `sf_polygon`.

@@

bw_growth_df <- bw_growth_df %>%
  filter(!foldID %in% c(19, 23, 21, 17, 8, 19)) %>%
  mutate(foldID = as.character(foldID))

# Deliverable
ggplot() +
  geom_sf(data = bw_growth_df, aes(col = foldID, alpha = buffer))

bw_cv_grid_sf <- bw_cv_grid$blocks %>%
  st_as_sf()
```

@@

### Focal versus competitor trees

@@

This requires the `growth_df` data frame; `max_dist`, the scalar defining competitive range; `cv_fold_size`, defining the size of the spatial cross-validation blocks; and the `id` variable as inputs.

```{r, eval = run_bw}
# Big Woods
if (!file.exists("focal_vs_comp_bw.Rdata")) {
  tic()
  focal_vs_comp_bw <- bw_growth_df %>% 
    create_focal_vs_comp(max_dist, cv_grid_sf = bw_cv_grid_sf, id = "treeID")
  toc()
  save(focal_vs_comp_bw, file = "focal_vs_comp_bw.Rdata")
} else {
  load("focal_vs_comp_bw.Rdata")
}
```

@@

Now we are ready to fit the competition model with `fit_bayesian_model`. This function needs only the `focal_vs_comp` as an input. Other options allow the user to specify prior parameters and run a species identity shuffle (see below).

```{r, eval = run_bw}
# Big Woods
tic()
posterior_param_bw <- focal_vs_comp_bw %>% 
  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
toc()
```

This output has the posterior parameters for the specified competition model. This `posterior_param` output can be used to get predicted growths for each individual (with `predict_bayesain_model`) to test how well the model performs. Or this `posterior_param` output can be plots (either the betas or lambdas) to understand what controls individual growth.

Here we calculate the RMSE

```{r, eval = run_bw}
# Big Woods
predictions <- focal_vs_comp_bw %>%
  predict_bayesian_model(posterior_param = posterior_param_bw) %>%
  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
predictions %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```

Now we test whether the identity of the competitor matters. We do this by shuffling the identity of competitors (but not of focal trees or spatial locations or sizes) and fitting the model again. We then compare RMSEs to see whether competitor identity matters to competitive effects

```{r, eval = run_bw}
posterior_param_bw_shuffle <- focal_vs_comp_bw %>%
  fit_bayesian_model(prior_param = NULL, run_shuffle = TRUE)

# b) Make predictions and compute RMSE
predictions_shuffle <- focal_vs_comp_bw %>%
  predict_bayesian_model(posterior_param = posterior_param_bw_shuffle) %>%
  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
predictions_shuffle %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```
